Bringing the Power of eBPF to Open vSwitch

### 2018-10-29 paper.tex
Minor fixes in ch1 and ch2

XDP provides another way for interacting with userspace programs, called AF\_XDP. 
AF\_XDP is a socket interface for control plane and a shared memory API for
accessing packets from userspace application. Using AF\_XDP, the OVS-AFXDP project
redirect packets to userspace, and processes the packets using OVS's full-fledged
userspace datapath implementation, dpif-netdev. In this approach, OVS-AFXDP
project treats the AF\_XDP as a fast packet-I/O channel.

@@

\subsection{OVS Forwarding Model}
OVS is widely used in virtualized data center environments as a software
switching layer inside various operating systems, including FreeBSD,
Windows Hyper-V, Solaris and Linux. As shown in
Figure~\ref{ovsintro}, the architecture of OVS consists of two major
components: a slow path and a fast path. OVS begins processing packets
in its datapath, the fast path, shortly after the packet is received by the NIC in the host OS.
The OVS datapath first performs packet parsing to extract relevant protocol
headers from the packet and store it locally in a manner that is efficient for
performing lookups (flow key), then it uses this information to look into the
match/action cache (flow table) and determines what needs to be done for this
packet. If there is no match in the flow table, the datapath passes the packet
from the kernel up to the slow path, \verb+ovs-vswitchd+, which maintains the
full determination of what needs to be executed to modify and forward
the packet correctly.  This process is called packet {\em upcall} and
usually happens at the first packet of a flow seen by the OVS datapath.
If the packet matches in this flow table, then the OVS datapath executes its
corresponding actions from the flow table lookup result and updates its flow
statistics.

@@

AF\_XDP was proposed and upstreamed to Linux kernel since 4.18~\cite{af_xdp_lwn}.
The core idea behind the AF\_XDP is to leverage the XDP eBPF program's
early access to the raw packet, and provide a high speed channel from the XDP
directly to a userspace socket interface. In other word, AF\_XDP socket family
connects the XDP packet receiving/sending path to the userspace, by-passing the
rest of the Linux networking stacks.
An AF\_XDP socket, called XSK, is created by using the normal socket() system
call. Unlike AF\_PACKET which uses the send() and receive() syscalls with packet
buffer as parameter, XSK introduces two rings in userspace: the Rx ring and the
Tx ring. These two rings are per-XSK data structure that the userspace program
needs to properly configure and maintain in order to receive and send packets.
Moreover, to provide zerocopy support for AF\_XDP, all the packet data buffers
used in Tx/Rx rings are allocated from a specific memory region called {\em umem}
which consists of a number of fixed size memory chunks.
Two rings, the Fill ring and the Completion ring, are associated with a umem,
and a umem can be shared between multiple XSKs.
Each element in the Rx/Tx ring or in Fill/Completion ring is a descriptor that
contains an address that points to a chunk in umem. The address is not system's
virtual or physical address but simply an offset within the entire umem memory
region.

For example, to receive packets from XSK, firstly, a set of descriptors pointing
to empty packet buffer need to be pushed into the Fill ring.  When a packet
arrives, kernel pops descriptors from the Fill ring, fills in the data into the
memory chunks pointed by the descriptors, and pushes the descriptors back to
the Rx ring.
The userspace program then checks the Rx ring, fetches the packet data from the
descriptors, and refills the empty buffer back to the Fill ring structure, so
that kernel can fill in new incoming packets later on.
For sending packets, the userspace program pushes a set of descriptors that
point to the packet buffers to the Tx ring, then issues sendmsg() system call.
Kernel consumes packet buffers asscoaited with the Tx ring, and pushes
transmitted descriptors to the Tx ring.
The userspace program then checks the Completion ring to determine whether the
packets have been sent.
In summary, XSK users need to properly program following four rings:
\begin{itemize}
\item Fill ring: for users to fill umem addresses to kernel for receiving packets.
\item Rx ring: for users to access received packets.
\item Tx ring: for users to place packets needed to be sent.
\item Completion ring: for users to check if packets are sent.
\end{itemize}
Unlike AF\_PACKET which is bound to entire netdev, the binding of XSK is more
fine-grained. XSK is bound to a specific queue on a device, so only the
traffic/flow sent to the queue shows up in the XSK.

>>>>

XDP provides another way for interacting with userspace programs, called AF\_XDP. 
AF\_XDP is a socket interface for control plane and a shared memory API for
accessing packets from userspace application. Using AF\_XDP, the OVS-AFXDP project
redirects packets to userspace, and processes the packets using OVS's full-fledged
userspace datapath implementation, dpif-netdev. In this approach, OVS-AFXDP
project treats the AF\_XDP as a fast packet-I/O channel.

@@

\subsection{OVS Forwarding Model}
OVS is widely used in virtualized data center environments as a software
switching layer inside various operating systems, including FreeBSD,
Windows Hyper-V, Solaris and Linux. As shown in
Figure~\ref{ovsintro}, the architecture of OVS consists of two major
components: a slow path and a fast path. OVS begins processing packets
in its datapath, the fast path, shortly after the packet is received by the NIC in the host OS.
The OVS datapath first performs packet parsing to extract relevant protocol
headers from the packet and stores it locally in a manner that is efficient for
performing lookups (flow key), then it uses this information to look into the
match/action cache (flow table) and determines what needs to be done for this
packet. If there is no match in the flow table, the datapath passes the packet
from the kernel up to the slow path, \verb+ovs-vswitchd+, which maintains the
full determination of what needs to be executed to modify and forward
the packet correctly.  This process is called packet {\em upcall} and
usually happens at the first packet of a flow seen by the OVS datapath.
If the packet matches in this flow table, then the OVS datapath executes its
corresponding actions from the flow table lookup result and updates its flow
statistics.

@@

AF\_XDP was proposed and upstreamed to Linux kernel since 4.18~\cite{af_xdp_lwn}.
The core idea behind the AF\_XDP is to leverage the XDP eBPF program's
early access to the raw packet, and provide a high speed channel from the XDP
directly to a userspace socket interface. In other word, AF\_XDP socket family
connects the XDP packet receiving/sending path to the userspace, by-passing the
rest of the Linux networking stacks.
An AF\_XDP socket, called XSK, is created by using the normal socket() system
call. Unlike AF\_PACKET which uses the send() and receive() syscalls with packet
buffer as parameter, XSK introduces two rings in userspace: the Rx ring and the
Tx ring. These two rings are per-XSK data structure that the userspace program
needs to properly configure and maintain in order to receive and send packets.
Moreover, to provide zerocopy support for AF\_XDP, all the packet data buffers
used in Tx/Rx rings are allocated from a specific memory region called {\em umem}
which consists of a number of fixed size memory chunks.
Two rings, the Fill ring and the Completion ring, are associated with a umem,
and a umem can be shared between multiple XSKs.
Each element in the Rx/Tx ring or in Fill/Completion ring is a descriptor that
contains an address that points to a chunk in umem. The address is not system's
virtual or physical address but simply an offset within the umem memory
region.

For example, to receive packets from XSK, firstly, the userspace program firstly
pushes a set of descriptors pointing to empty packet buffer into the Fill ring.
When a packet
arrives, kernel pops descriptors from the Fill ring, fills in the data into the
memory chunks pointed by the descriptors, and pushes the descriptors back to
the Rx ring.
The userspace program then checks the Rx ring, fetches the packet data from the
descriptors, and refills the empty buffer back to the Fill ring structure, so
that kernel can fill in new incoming packets later on.
For sending packets, the userspace program pushes a set of descriptors that
point to the packet buffers to the Tx ring, then issues sendmsg() system call.
Kernel consumes packet buffers asscoaited with the Tx ring, and pushes
transmitted descriptors to the Tx ring.
The userspace program then checks the Completion ring to determine whether the
packets have been sent.
In summary, XSK users need to properly program following four rings:
\begin{itemize}
\item Fill ring: for users to fill umem addresses to kernel for receiving packets.
\item Rx ring: for users to access received packets.
\item Tx ring: for users to place packets needed to be sent.
\item Completion ring: for users to check if packets are sent.
\end{itemize}
Unlike AF\_PACKET which is bound to entire netdev, the binding of XSK is more
fine-grained. XSK is bound to a specific queue on a device, so only the
traffic sent to the queue shows up in the XSK.
